---
title: "Introduction of efdcr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction of efdcr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is efdcr
The `efdcr` package is designed to faciliate the modelling process with the EFDC (Environment Fluid Dynamic Code) and [EE modeling system (EEMS)](https://www.eemodelingsystem.com/).
It provides functions for pre- and post-processing the data reqired and generated by the EEMS.
Note that the package is currently under development, the APIs will be changed with more functionalities being added. I will try my best to keep the consistency of the APIs and to update carefully. 

## Installation of efdcr
You can Install the latest development version of efdcr from [GitHub](https://www.github.com) with:

```{r,eval=FALSE}
if (!require("devtools")) {
  install.packages("devtools")
}
devtools::install_github("hxfan1227/efdcr")
```

## Work with efdcr
To work with `efdcr`, you should have some knowlege about the 'tidy data' format, details of which can be found at <https://www.jstatsoft.org/article/view/v059i10>.


